Simulating High-Dimensional Multivariate Data

### 2020-12-29 030-bigsimr-package.Rmd

```{r ch030-cor-bounds, echo=TRUE}
cor_bounds(margins = margins, type = type)
```

@@

```{r ch030-cor-bounds-pearson, echo=TRUE}
cor_bounds(margins = margins, type = 'pearson', reps = 1e5)
```

@@

```{r ch030-sim-margins, echo=TRUE}
x <- rvec(10000, R_S, margins, type)
df_sim <- as.data.frame(x)
colnames(df_sim) <- colnames(df)
```

>>>>

```{r ch030-cor-bounds, echo=TRUE}
cor_bounds(margins = margins, method = type)
```

@@

```{r ch030-cor-bounds-pearson, echo=TRUE}
cor_bounds(margins = margins, method = 'pearson', reps = 1e6)
```

@@

```{r ch030-sim-margins, echo=TRUE}
x <- rvec(10000, rho, margins)
df_sim <- as.data.frame(x)
colnames(df_sim) <- colnames(df)
```

### 2020-12-29 040-monte-carlo-evaluation.Rmd
4th draft. Address Ed comments Part I.

```{r ch040-biNormal, echo = FALSE, eval = TRUE}
mu <- 0
sigma <- 1
if ( !file.exists( 'results/biNorm_sims.rds' ) )  {
    mom_norm <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(mean = m, sd = s)
    }
    mu <- 0
    sigma <- 1
    margins <- alist(
        qnorm(mean = mu, sd = sigma),
        qnorm(mean = mu, sd = sigma)
    )
    type <- c("pearson", "spearman", "kendall")
    cores <- as.integer( parallel::detectCores() - 1 )
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)
    adjustForDiscrete <- c(FALSE)
    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = type, cores = cores, n = n,
                            stringsAsFactors = FALSE,
                            adjustForDiscrete = adjustForDiscrete)
    res <- data.frame()
    dir.create("./results/biNorm_sims", showWarnings = TRUE, recursive = TRUE)
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
        tmp_bounds <- cor_bounds(margins, type = type)
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)
        ## rho = cor_seq[1]
        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0
            ## Rho <- convertCor( rho = Rho, from = type, to = 'pearson' )
            time_data <- system.time({
                ## x <- rmvn(n = n, mu = rep( mu, 2 ), sigma = Rho)
                ## use rvec to test our software
                x <- rvec(n = n,
                          rho = Rho,
                          margins = margins,
                          type = type,
                          cores = cores)
            })
            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-adj", as.character(adjustForDiscrete),
                "-dev", "1CORE",
                "-lib", "bigsimr"
            )
            ## save sims in their own subdirectory
            saveRDS(x, file = paste0("./results/biNorm_sims/", id, ".rds"))
            ## Estimate statistics
            Rho_hat <- cor_fast(x, method = type)
            rho_hat <- Rho_hat[1, 2]
            norm_args_hat <- mom_norm(x[,1])
            mu_hat <- norm_args_hat$mean
            sigma_hat <- norm_args_hat$sd
            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "norm",
                                  adjustForDiscrete = adjustForDiscrete,
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  mean = mu,
                                  sd = sigma,
                                  mean_hat = mu_hat,
                                  sd_hat = sigma_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biNorm_sims.rds")
}
```

@@

        tmp_bounds <- cor_bounds(margins, type = type)
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

@@

    dir.create("./results/biNB_sims", showWarnings = TRUE, recursive = TRUE)
    res <- data.frame()
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
        ## AGS. 17 Oct 2020. cor_bounds failed.
        ## I'll create a minimal working example and open an issue
        ## work around for now
        if ( type == 'pearson' ) {
            tmp_bounds <- cor_bounds(margins, type = type)
            cor_lo <- tmp_bounds$lower[1,2] + eps
            cor_hi <- tmp_bounds$upper[1,2] - eps
        } else {
            cor_lo <- -1 + eps
            cor_hi <- 1 - eps
        }
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

@@

## AGS 12/22/20: Removing the latex formating fixes the html render.
## But how to improve latex table formating? Better to use the template
knitr::kable(tabMAE,
             ## format = 'latex',
             col.names = c("No. of random vectors",
                           "Correlation type",
                           "Distribution",
                           "Mean abs. error"),
             ## booktabs = TRUE,
             caption = 'Average abolute error in matching the target dependency across the entire range of possible correlations for each bivariate marginal.',
             ## linesep = c('', '',  '\\addlinespace')
)
```

>>>>

```{r ch040-biNormal, echo = FALSE, eval = TRUE}
mu <- 0
sigma <- 1
if ( !file.exists( 'results/biNorm_sims.rds' ) )  {
    mom_norm <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(mean = m, sd = s)
    }
    mu <- 0
    sigma <- 1
    margins <- alist(
        qnorm(mean = mu, sd = sigma),
        qnorm(mean = mu, sd = sigma)
    )
    type <- c("pearson", "spearman", "kendall")
    cores <- as.integer( parallel::detectCores() - 1 )
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)
    adjustForDiscrete <- c(FALSE)
    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = type, cores = cores, n = n,
                            stringsAsFactors = FALSE,
                            adjustForDiscrete = adjustForDiscrete)
    res <- data.frame()
    dir.create("./results/biNorm_sims", showWarnings = TRUE, recursive = TRUE)
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
        tmp_bounds <- cor_bounds(margins, method = type)
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)
        ## rho = cor_seq[1]
        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0
            ## Rho <- convertCor( rho = Rho, from = type, to = 'pearson' )
            time_data <- system.time({
                ## x <- rmvn(n = n, mu = rep( mu, 2 ), sigma = Rho)
                ## use rvec to test our software
                x <- rvec(n = n,
                          rho = Rho,
                          margins = margins,
                          type = type,
                          cores = cores)
            })
            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-adj", as.character(adjustForDiscrete),
                "-dev", "1CORE",
                "-lib", "bigsimr"
            )
            ## save sims in their own subdirectory
            saveRDS(x, file = paste0("./results/biNorm_sims/", id, ".rds"))
            ## Estimate statistics
            Rho_hat <- cor_fast(x, method = type)
            rho_hat <- Rho_hat[1, 2]
            norm_args_hat <- mom_norm(x[,1])
            mu_hat <- norm_args_hat$mean
            sigma_hat <- norm_args_hat$sd
            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "norm",
                                  adjustForDiscrete = adjustForDiscrete,
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  mean = mu,
                                  sd = sigma,
                                  mean_hat = mu_hat,
                                  sd_hat = sigma_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biNorm_sims.rds")
}
```

@@

        tmp_bounds <- cor_bounds(margins, method = type)
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

@@

    dir.create("./results/biNB_sims", showWarnings = TRUE, recursive = TRUE)
    res <- data.frame()
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
        ## AGS. 17 Oct 2020. cor_bounds failed.
        ## I'll create a minimal working example and open an issue
        ## work around for now
        if ( type == 'pearson' ) {
            tmp_bounds <- cor_bounds(margins, method = type)
            cor_lo <- tmp_bounds$lower[1,2] + eps
            cor_hi <- tmp_bounds$upper[1,2] - eps
        } else {
            cor_lo <- -1 + eps
            cor_hi <- 1 - eps
        }
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

@@

## AGS 12/22/20: Removing the latex formatting fixes the html render.
## But how to improve latex table formatting? Better to use the template
knitr::kable(tabMAE,
             ## format = 'latex',
             col.names = c("No. of random vectors",
                           "Correlation type",
                           "Distribution",
                           "Mean abs. error"),
             ## booktabs = TRUE,
             caption = 'Average abolute error in matching the target dependency across the entire range of possible correlations for each bivariate marginal.',
             ## linesep = c('', '',  '\\addlinespace')
)
```
